---
title: "exclude outliers and recalculate AC, AN, AF"
output: html_document
editor_options: 
  chunk_output_type: console
---

started: AL26Apr2019  
last updated: AL01Oct2019  

# Summary  

Exclude 34 eigenvector outliers detected in joined Ampliseq-1KG PCA  

Then:  
- exclude variants that were present in the excluded samples only  
- recalculate AC-AN-AF-s in whole data and in UBC and CBC sub-groups  

Input:  
- 3,675 vars x 515 samples (258UBC + 257CBC)  

Output:  
- 3,044 vars x 481 samples (240UBC + 241CBC)  

# Start_section

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

```{r set_width, echo=F}

#eval=F

options(width = 999)
# https://stackoverflow.com/questions/36845178/width-of-r-code-chunk-output-in-rmarkdown-files-knitr-ed-to-html

```

```{r start_section}

Sys.time()
rm(list=ls())
graphics.off()

library(knitr)
library(dplyr)

base_folder="/Users/alexey/Documents/wecare/ampliseq/v05_ampliseq_only/s13_ampliseq_only_PCA/s03_explore_PCA_plots_exclude_outliers"

opts_knit$set(root.dir = base_folder)

options(stringsAsFactors = F)
options(warnPartialMatchArgs = T, 
        warnPartialMatchAttr = T, 
        warnPartialMatchDollar = T)

#options(error = browser()) # Type Q or c to exit, drop browser level
# https://support.rstudio.com/hc/en-us/articles/200713843?version=1.1.456&mode=desktop
# https://stackoverflow.com/questions/13052522/how-to-leave-the-r-browser-mode-in-the-console-window/13052588 

```

# Read_data  

```{r read_data}

load(paste(base_folder, "s02_explore_ethnicity_of_outliers.RData", sep="/"))

```

# Check data

```{r}

ls()

dim(genotypes.mx)
dim(phenotypes.df)
dim(variants.df)

sum(colnames(genotypes.mx) != rownames(phenotypes.df))
sum(rownames(genotypes.mx) != rownames(variants.df))

```

# Remove 34 PCA outliers

Also  

- remove unnecesary columns from phenotypes data frame  
- sync genotypes matrix  

```{r remove_outliers}

# Remove outliers
phenotypes.df <- phenotypes.df[!phenotypes.df$joined_pca_outlier,]

# Remove unnecessary columns
colnames(phenotypes.df)
phenotypes.df <- phenotypes.df[,c(1:31,37)]
colnames(phenotypes.df)

# Syncronise genotypes
genotypes.mx <- genotypes.mx[,rownames(phenotypes.df)]

# Check result
dim(genotypes.mx)
dim(phenotypes.df)
table(phenotypes.df$cc)

```

# Update variants

```{r}

# Function to detect uniform vectors (expects vectors containing 0,1,2 or NA only)
uniform.udf <- function(x){
  if(all(is.na(x))) return(TRUE) # Return T if all are NA
  min(x,na.rm=T)==max(x,na.rm=T)} # Return T if min=max, otherwise return F

uniform_variants <- apply(genotypes.mx, 1, uniform.udf)
sum(uniform_variants)

genotypes.mx <- genotypes.mx[!uniform_variants,]
variants.df <- variants.df[!uniform_variants,]

dim(genotypes.mx)
dim(variants.df)

rm(uniform_variants, uniform.udf)

```

# Recalculate AC-AN-AF-s

## AN function and groups check

In general, this simpified version of AN function should only work for autosomes;   
However, it works here for X chromosome too, because all the subjects are females.  

```{r an_and_counts}

an.udf <- function(x){2*sum(!is.na(x))}
summary(as.factor(phenotypes.df$group))

```

## Total AC-AN-AF

```{r total_acanaf}

total_AC <- apply(genotypes.mx, 1, sum, na.rm=T)
total_AN <- apply(genotypes.mx, 1, an.udf)
total_AF <- total_AC/total_AN
hist(total_AF, lab=T, ylim=c(0,3000))

```

## UBC AC-AN-AF

```{r ubc_acanaf}

ubc_genotypes.mx <- genotypes.mx[,phenotypes.df$group=="UBC"]
dim(ubc_genotypes.mx)

ubc_AC <- apply(ubc_genotypes.mx, 1, sum, na.rm=T)
ubc_AN <- apply(ubc_genotypes.mx, 1, an.udf)
ubc_AF <- ubc_AC/ubc_AN
hist(ubc_AF, lab=T, ylim=c(0,3000))

rm(ubc_genotypes.mx)

```

## CBC AC-AN-AF

```{r cbc_acanaf}

cbc_genotypes.mx <- genotypes.mx[,phenotypes.df$group=="CBC"]
dim(cbc_genotypes.mx)

cbc_AC <- apply(cbc_genotypes.mx, 1, sum, na.rm=T)
cbc_AN <- apply(cbc_genotypes.mx, 1, an.udf)
cbc_AF <- cbc_AC/cbc_AN
hist(cbc_AF, lab=T, ylim=c(0,3000))

rm(cbc_genotypes.mx, an.udf)

```

## Add acanaf-s to variants table

```{r update_variants}

variants.df <- data.frame(variants.df,
  total_AC, total_AN, total_AF,
  ubc_AC, ubc_AN, ubc_AF,
  cbc_AC, cbc_AN, cbc_AF)

colnames(variants.df)

rm(total_AC, total_AN, total_AF,
  ubc_AC, ubc_AN, ubc_AF,
  cbc_AC, cbc_AN, cbc_AF)

```

# Check data

```{r}

ls()

dim(genotypes.mx)
dim(phenotypes.df)
dim(variants.df)

sum(colnames(genotypes.mx) != rownames(phenotypes.df))
sum(rownames(genotypes.mx) != rownames(variants.df))

```

# Save results

```{r save_results}

save.image(paste(base_folder, "s03_exclude_outliers_recalculate_acanaf.RData", sep="/"))

```

# Final_section

```{r final_section}

ls()
sessionInfo()
Sys.time()

```
